---
output: github_document
encoding: 'UTF-8'
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

  <!-- badges: start -->
  [![R-CMD-check](https://github.com/xmarquez/democracyData/workflows/R-CMD-check/badge.svg)](https://github.com/xmarquez/democracyData/actions)
  <!-- badges: end -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# democracyData

This package archives a large number of datasets measuring democracy in use in the scholarly literature, and it provides functions to access many others. You can use it to access some widely used datasets, including [Polity5](http://www.systemicpeace.org/inscrdata.html), [Freedom House](https://freedomhouse.org/), [Geddes, Wright, and Frantz' autocratic regimes dataset](http://sites.psu.edu/dictators/), the [Lexical Index of Electoral Democracy](https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/29106), the [DD/ACLP/PACL/CGV dataset](https://sites.google.com/site/joseantoniocheibub/datasets/democracy-and-dictatorship-revisited), the main indexes of the [V-Dem dataset](https://www.v-dem.net/vdemds.html), and many others.

## Installation

The package is only available on Github. Install as follows:

```
remotes::install_github("xmarquez/democracyData")

```

## Basic usage

For the vast majority of use cases, you can just type the name of the dataset you require. For example, here's the [DD/ACLP/PACL/CGV](https://sites.google.com/site/joseantoniocheibub/datasets/democracy-and-dictatorship-revisited) dataset:

```{r}
library(democracyData)
pacl
```

Here's [Polity IV](http://www.systemicpeace.org/inscrdata.html):

```{r}
polityIV
```

And here's a basic version of the V-Dem dataset, including only the 7 main indexes of democracy:

```{r}
vdem_simple
```

All datasets in this package are fully documented; type `?pacl` for example to see the documentation for the PACL dataset.

## Downloading democracy data

Three democracy datasets are not currently archived in this package: the family of datasets released by [Freedom House](https://freedomhouse.org/); the [Polity5](http://www.systemicpeace.org/inscrdata.html) dataset (which superceded [Polity IV](http://www.systemicpeace.org/inscrdata.html)); and the Voice and Accountability indexes of the [World Governance Indicators](http://info.worldbank.org/governance/wgi/). (The full V-Dem dataset is also not archived here; you can access it via the [vdemdata](https://github.com/vdeminstitute/vdemdata) package). To download these datasets, use the the `download_*` family of functions.

For example, we can download and process the Freedom House "Freedom in the World" dataset as follows:

```{r, message = FALSE}
fh <- download_fh(verbose = FALSE)

fh 

```

This downloads the latest update of the "Freedom in the World" dataset (1972-2021, corresponding to the 2022 report), puts it in country-year format (extracting the relevant info from the awful Excel table that Freedom House makes available), calculates the variables `fh_total` and `fh_total_reversed`, and adds state system information, including a standardized country name, the [Gleditsch-Ward](http://privatewww.essex.ac.uk/~ksg/statelist.html) country code and the [Correlates of War](http://www.correlatesofwar.org/data-sets/state-system-membership) country code. 

Other democracy datasets included in this package do not need to be downloaded, but they can often also be "re-downloaded" from the websites of their creators or maintainers if required. For example, one can either access PACL directly by typing

```{r}
pacl
```

Or re-download the dataset from Jose Antonio Cheibub's website as follows:

```{r}

pacl_redownloaded <- redownload_pacl(verbose = FALSE)

pacl_redownloaded

```

These two data frames should be identical:

```{r}

identical(pacl, pacl_redownloaded)

```

You should thus normally use the "archived" versions of these datasets, unless you want to manipulate the raw data yourself (using the `redownload_*` functions with the option `return_raw = TRUE`), or think they might have been updated since you installed this package. 

## Included democracy datasets

For a list of all the democracy datasets available through this package, type `democracy_info`:

```{r}
library(dplyr)

democracy_info %>%
  knitr::kable()

```

## Combining all democracy datasets

You can create one huge data frame including all democracy measures with one call:

```{r, message = FALSE}

democracy_data <- generate_democracy_scores_dataset(output_format = "wide",
                                                    verbose = FALSE)

democracy_data
```

This can take some time, since it downloads all downloadable datasets (Freedom House, Polity 5, and the WGI Voice and Accountability index), processes them (adds state system information, puts them in country-year format, fixes wrong codes, etc.), and matches them to all the other datasets. In any case, you can select exactly which datasets to include in your big data frame. See `?generate_democracy_scores_dataset` for further options to customize the output.


## Latent Variable Indexes of Democracy

The package also offers a series of convenience functions to calculate latent variable indexes of democracy (following Pemstein, Meserve, and Melton's [2010 article](https://www.jstor.org/stable/25792023?seq=1) "Democratic Compromise: A Latent Variable Analysis of Ten Measures of Regime Type"); see the vignette on [Replicating and Extending the UD scores of Pemstein, Meserve, and Melton](https://xmarquez.github.io/democracyData/articles/Replicating_and_extending_the_UD_scores.html). It also contains a pre-calculated extended version of these scores, available as `extended_uds`:

```{r}
extended_uds
```


## State system functions

The package also includes a couple of other convenience functions to work with historical democracy data and determine state system membership. The first is `country_year_coder`, which works like the [`countrycode`](https://CRAN.R-project.org/package=countrycode) package, except that it is able to determine state system information for country-year pairs. Suppose you have this dataset:

```{r}
my_weird_democracy_data <- tibble(country = c("Germany", "Germany", "Germany", "Germany", "East Germany",
                                    "Federal Republic of Germany",
                                    "Somaliland", "Somalia",
                                    "Palestine", "Russia",
                                    "Russia", "USSR",
                                    "Republic of Vietnam",
                                    "Yugoslavia", 'Yugoslavia',
                                    "Vietnam, South"),
                        year = c( 2015, 1930, 1970, 1945, 1949,
                                 1992, 1990,
                                 1990, 1940,
                                 1917, 1912,
                                 1922, 1975,
                                 1990, 1991, 1954),
                        my_measure = rnorm(16))


my_weird_democracy_data
```

and you then want to add state system information. `country_year_coder` does that for you!

```{r}

my_weird_democracy_data <- my_weird_democracy_data %>%
  country_year_coder(country,
                     year,
                     match_type = "country",
                     verbose = FALSE,
                     include_in_output = c("extended_country_name", 
                                           "GWn", "cown", 
                                           "polity_ccode", 
                                           "in_GW_system", 
                                           "in_cow_system", 
                                           "in_polity_system",
                                           "polity_startdate",
                                           "polity_enddate"))

my_weird_democracy_data %>%
  knitr::kable()
```

`country_year_coder` tries to match not just the country name or the country code (as `countrycode` does), but also to figure out the appropriate state system code *given* the year. (Above, for example, the function figures out that Germany 1970 should get a COW code of 260, but Germany 1992 should get 255 - though it should retain the 260 code in the Gleditsch and Ward system of states. This is, incidentally, how `download_fh` adds the correct COW and GW country codes to Freedom House's Excel data). It also tries to figure out whether a given country-year is in the specific state system list. (In the example above, Germany in 1945 is not listed as a member of the state system in either COW or Gleditsch and Ward, since it was occupied by the Allies as of 31 December 1945, but is listed as a member of the state system in Polity IV as the Federal Republic, though with a polity score of -66, "interregnum").

One nice thing about `country_year_coder` (in my humble opinion!) is that it can sometimes correct country coding errors; I've run across more than one dataset with the supposed COW code 255 for the Federal Republic of Germany for the period 1955-1990, which would prevent a clean join to a dataset with the correct COW code, but would be caught by `country_year_coder`.

There is also a function that allows you to create a blank state system panel for any of the three main state systems:

```{r}
create_panel(system = "cow")

create_panel(system = "GW")
```

## Citation

The standard `citation` function from base `R` will produce a list of citations for all the datasets included in this package:

```{r, results='asis'}
citation(package = "democracyData")
```

You can also find the citation for a specific dataset using the wrapper `cite_dataset` with the name of the dataset in this package:

```{r, results='asis'}
cite_dataset("gwf")
```

# Feedback and Caveats

Feedback welcome! 

Note that some functions in this package can be quite slow: generating a full democracy dataset (including downloading Freedom House, Polity, and WGI) or applying `country_year_coder` to a large data frame both can take some time. Suggestions to accelerate the code are welcome. 

`country_year_coder` fails to give correct answers in some weird edge cases mostly involving Yugoslavia, Germany, or Vietnam. If you run across any of these cases, let me know.
