---
title: "Adding all democracy datasets, testing Targets pipeline works"
author: "Xavier Marquez"
date: "`r lubridate::now()`"
output: 
  html_document: 
    toc: yes
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(echo = TRUE,
                      warning=TRUE,
                      message = TRUE,
                      error = FALSE)
library(tidyverse)
library(democracyData)
library(targets)

verbose <- TRUE
include_in_output <-  c("extended_country_name", "GWn", "cown", "in_GW_system")
pmm_replication <-  prepare_pmm_replication_data(path = here::here("data-raw/democracy1946.2008.rda"),
                                                 verbose = verbose,
                                                 include_in_output = include_in_output) 
```

## Anckar and Fredriksson

```{r anckar, results = 'asis'}
cite_dataset("anckar")

anckar <- redownload_anckar(verbose = verbose,
                            include_in_output = include_in_output)

stopifnot(identical(anckar,tar_read(anckar, store = here::here("_targets"))))
stopifnot(identical(anckar, democracyData::anckar))
```

## ANRR

```{r anrr, results = 'asis'}

cite_dataset("anrr")

anrr <- prepare_anrr(path = here::here("data-raw/DDCGdata_final.dta"),
                     verbose = verbose,
                     include_in_output = include_in_output)

stopifnot(identical(anrr,tar_read(anrr, store = here::here("_targets"))))
stopifnot(identical(anrr, democracyData::anrr))
```

## Arat

```{r arat, results = 'asis'}
cite_dataset("arat")

arat_pmm <- extract_pmm_var(pmm_replication, pmm_arat, 
                            include_in_output = include_in_output)

stopifnot(identical(arat_pmm,tar_read(arat_pmm, store = here::here("_targets"))))
stopifnot(identical(arat_pmm, democracyData::arat_pmm))
```

## BLM

```{r blm, results = 'asis'}
cite_dataset("blm")

# Expired certificate
# blm <- redownload_blm(verbose = verbose,
#                       include_in_output = include_in_output)

blm_pmm <- extract_pmm_var(pmm_replication, pmm_blm,
                           include_in_output = include_in_output)

# usethis::use_data(blm, overwrite = TRUE)

stopifnot(identical(blm_pmm,tar_read(blm_pmm, store = here::here("_targets"))))
stopifnot(identical(blm_pmm, democracyData::blm_pmm))
```

# BMR (Boix Miller Rosato)

```{r bmr, results = 'asis'}
cite_dataset("bmr")

bmr <- redownload_bmr(verbose = verbose, 
                      include_in_output = include_in_output)

stopifnot(identical(bmr,tar_read(bmr, store = here::here("_targets"))))
stopifnot(identical(bmr, democracyData::bmr))
```

# BNR

Redownload not fixable - 

```{r bnr, results = 'asis'}
cite_dataset("bnr")

# bnr <- redownload_bnr(include_in_output = include_in_output, extend = FALSE)
# 
# bnr_extended <- redownload_bnr(include_in_output = include_in_output, extend = TRUE)
# 
# anti_join(bnr_extended, bnr %>% select(-bnr_country)) %>%
#         group_by(bnr_country) %>%
#         arrange(bnr_country, cown, year) %>%
#         mutate(period = count_sequence_breaks(year)) %>%
#         group_by(bnr_country, bnr, period) %>%
#         summarise(min= min(year), max = max(year)) %>%
#   knitr::kable()
# 
# anti_join(bnr %>% select(-bnr_country), bnr_extended) %>%
#   knitr::kable()


```

# BTI

```{r bti, results = 'asis'}
cite_dataset("bti")

bti <- redownload_bti(verbose = verbose, 
                      include_in_output = include_in_output)

stopifnot(identical(bti, tar_read(bti, store = here::here("_targets"))))
stopifnot(identical(bti, democracyData::bti))

```

# Bollen

```{r bollen, results = 'asis'}
cite_dataset("bollen")

bollen_pmm <- extract_pmm_var(pmm_replication, pmm_bollen,
                              include_in_output = include_in_output)

stopifnot(identical(bollen_pmm, tar_read(bollen_pmm, store = here::here("_targets"))))
stopifnot(identical(bollen_pmm, democracyData::bollen_pmm))
```

# Doorenspleet

```{r doorenspleet, results = 'asis'}
cite_dataset("doorenspleet")

doorenspleet <- prepare_doorenspleet(path = here::here("data-raw/Doorenspleet data.csv"),
                                     verbose = verbose, 
                                     include_in_output = include_in_output)

stopifnot(identical(doorenspleet, tar_read(doorenspleet, store = here::here("_targets"))))
stopifnot(identical(doorenspleet, democracyData::doorenspleet))

```

# WGI

```{r wgi, results = 'asis'}
cite_dataset("wgi")

wgi <- download_wgi_voice_and_accountability(verbose = verbose, 
                                             include_in_output = include_in_output)

stopifnot(identical(wgi, tar_read(wgi, store = here::here("_targets"))))
stopifnot(identical(wgi, democracyData::wgi))

```

# EIU

```{r eiu, results= 'asis'}
cite_dataset("eiu")

eiu <- prepare_eiu(path = here::here("data-raw/DI-final-version-report-2022.pdf"),
                   verbose = verbose, include_in_output = include_in_output)

stopifnot(identical(eiu, tar_read(eiu, store = here::here("_targets"))))
stopifnot(identical(eiu, democracyData::eiu))
```

# Freedom House

```{r fh, results = 'asis'}
cite_dataset("fh")

fh <- download_fh(verbose = verbose,
                  include_in_output = include_in_output)

fh_with_territories <- download_fh(verbose = verbose,
                                   include_in_output = include_in_output,
                                   include_territories = TRUE)

fh_electoral <- download_fh_electoral(verbose = verbose,
                                      include_in_output = include_in_output)

fh_full <- download_fh_full(verbose = verbose,
                                      include_in_output = include_in_output)

fh_pmm <- extract_pmm_var(pmm_replication, pmm_freedomhouse,
                          include_in_output = include_in_output) 

stopifnot(identical(fh, tar_read(fh, store = here::here("_targets"))))
stopifnot(identical(fh_with_territories, tar_read(fh_with_territories, store = here::here("_targets"))))
stopifnot(identical(fh_electoral, tar_read(fh_electoral, store = here::here("_targets"))))
stopifnot(identical(fh_full, tar_read(fh_full, store = here::here("_targets"))))
stopifnot(identical(fh_pmm, tar_read(fh_pmm, store = here::here("_targets"))))
stopifnot(identical(fh_pmm, democracyData::fh_pmm))
```

# GWF

```{r gwf, results = 'asis'}
cite_dataset("gwf")

gwf_autocratic <- redownload_gwf(verbose = verbose,
                                 dataset = "autocratic",
                                 include_in_output = include_in_output)

gwf_autocratic_extended <- redownload_gwf(verbose = verbose,
                                          dataset = "autocratic",
                                          include_in_output = include_in_output,
                                          extend = TRUE)

gwf_all <- redownload_gwf(verbose = verbose,
                          dataset = "all",
                          include_in_output = include_in_output)

gwf_all_extended <- redownload_gwf(verbose = verbose,
                                   dataset = "all",
                                   include_in_output = include_in_output,
                                   extend = TRUE)

stopifnot(identical(gwf_autocratic, tar_read(gwf_autocratic, store = here::here("_targets"))))
stopifnot(identical(gwf_autocratic, democracyData::gwf_autocratic))
stopifnot(identical(gwf_autocratic_extended, tar_read(gwf_autocratic_extended, store = here::here("_targets"))))
stopifnot(identical(gwf_autocratic_extended, democracyData::gwf_autocratic_extended))
stopifnot(identical(gwf_all, tar_read(gwf_all, store = here::here("_targets"))))
stopifnot(identical(gwf_all, democracyData::gwf_all))
stopifnot(identical(gwf_all_extended, tar_read(gwf_all_extended, store = here::here("_targets"))))
stopifnot(identical(gwf_all_extended, democracyData::gwf_all_extended))

```

# Hadenius

```{r hadenius, results = 'asis'}
cite_dataset("hadenius")

hadenius_pmm <- extract_pmm_var(pmm_replication, pmm_hadenius,
                                include_in_output = include_in_output)

stopifnot(identical(hadenius_pmm, tar_read(hadenius_pmm, store = here::here("_targets"))))
stopifnot(identical(hadenius_pmm, democracyData::hadenius_pmm))

```

# Kailitz

```{r kailitz, results = 'asis'}
cite_dataset("kailitz")

kailitz <- prepare_kailitz(path = here::here("data-raw/kailitz.yearly.rds"),
                           verbose = verbose, include_in_output = include_in_output)

stopifnot(identical(kailitz, tar_read(kailitz, store = here::here("_targets"))))
stopifnot(identical(kailitz, democracyData::kailitz))
```


# LIED

```{r lied, results = 'asis'}
cite_dataset("LIED")
      
LIED <- redownload_lied(verbose = verbose,
                        include_in_output = include_in_output)

stopifnot(identical(LIED, tar_read(LIED, store = here::here("_targets"))))
stopifnot(identical(LIED, democracyData::LIED))
```


# Magaloni

```{r magaloni, results = 'asis'}
cite_dataset("magaloni")

magaloni <- redownload_magaloni(verbose = verbose, 
                                include_in_output = include_in_output)

magaloni_extended <- redownload_magaloni(extend = TRUE,
                                         verbose = verbose, 
                                         include_in_output = include_in_output)

anti_join(magaloni_extended, magaloni) %>%
    group_by(magaloni_country, magaloni_ccode, cown) %>% summarise(min(year), max(year), n()) %>%
  knitr::kable()


anti_join(magaloni_extended, magaloni) %>%
  filter(grepl("Pakis",magaloni_country)) %>%
  knitr::kable()

stopifnot(identical(magaloni, tar_read(magaloni, store = here::here("_targets"))))
stopifnot(identical(magaloni, democracyData::magaloni))
stopifnot(identical(magaloni_extended, tar_read(magaloni_extended, store = here::here("_targets"))))
stopifnot(identical(magaloni_extended, democracyData::magaloni_extended))
```


# Mainwaring

```{r mainwaring, results = 'asis'}
cite_dataset("mainwaring")

mainwaring <- prepare_mainwaring(path = here::here("data-raw/Mainwaring Linan.txt"), 
                                 verbose = verbose,
                                 include_in_output = include_in_output)

mainwaring_pmm <- extract_pmm_var(pmm_replication, pmm_mainwaring,
                                 include_in_output = include_in_output)

stopifnot(identical(mainwaring, tar_read(mainwaring, store = here::here("_targets"))))
stopifnot(identical(mainwaring, democracyData::mainwaring))
stopifnot(identical(mainwaring_pmm, tar_read(mainwaring_pmm, store = here::here("_targets"))))
stopifnot(identical(mainwaring_pmm, democracyData::mainwaring_pmm))

```

# Munck

```{r munck, results = 'asis'}

cite_dataset("munck")

munck_pmm <- extract_pmm_var(pmm_replication, pmm_munck,
                             include_in_output = include_in_output)

stopifnot(identical(munck_pmm, tar_read(munck_pmm, store = here::here("_targets"))))
stopifnot(identical(munck_pmm, democracyData::munck_pmm))
```

# PACL

```{r pacl, results = 'asis'}
cite_dataset("pacl")

pacl_pmm <- extract_pmm_var(pmm_replication, pmm_pacl,
                            include_in_output = include_in_output)

pacl <- redownload_pacl(verbose = verbose,
                        include_in_output = include_in_output)

stopifnot(identical(pacl_pmm, tar_read(pacl_pmm, store = here::here("_targets"))))
stopifnot(identical(pacl_pmm, democracyData::pacl_pmm))
stopifnot(identical(pacl, tar_read(pacl, store = here::here("_targets"))))
stopifnot(identical(pacl, democracyData::pacl))
```
# PACL Update

```{r paclUpdate, results = 'asis'}
cite_dataset("pacl_update")

pacl_update <- redownload_pacl_update(verbose = verbose,
                                      include_in_output = include_in_output)

stopifnot(identical(pacl_update, tar_read(pacl_update, store = here::here("_targets"))))
stopifnot(identical(pacl_update, democracyData::pacl_update))
```

# PEPS 

```{r peps, results = 'asis'}
cite_dataset("peps")

peps <- redownload_peps(verbose = verbose,
                        include_in_output = include_in_output)

stopifnot(identical(peps, tar_read(peps, store = here::here("_targets"))))
stopifnot(identical(peps, democracyData::peps))
```

# Polity

```{r polity, results = 'asis'}
cite_dataset("polity")

polity_pmm <- extract_pmm_var(pmm_replication, pmm_polity,
                              include_in_output = include_in_output)

polityIV <- redownload_polityIV(verbose = verbose,
                                 include_in_output = include_in_output)

polity5 <- download_polity_annual(verbose = verbose,
                                 include_in_output = include_in_output)

stopifnot(identical(polity_pmm, tar_read(polity_pmm, store = here::here("_targets"))))
stopifnot(identical(polity_pmm, democracyData::polity_pmm))
stopifnot(identical(polityIV, tar_read(polityIV, store = here::here("_targets"))))
stopifnot(identical(polityIV, democracyData::polityIV))
stopifnot(identical(polity5, tar_read(polity5)))
stopifnot(identical(polity5, democracyData::polity5))
```

# PITF

```{r pitf, results = 'asis'}
cite_dataset("pitf")

pitf_p4 <- create_pitf_scores(polityIV, verbose = verbose, include_in_output = include_in_output)

pitf <- create_pitf_scores(verbose = verbose, include_in_output = include_in_output)

stopifnot(identical(pitf_p4, tar_read(pitf_p4)))
stopifnot(identical(pitf_p4, democracyData::pitf_p4))
stopifnot(identical(pitf, tar_read(pitf, store = here::here("_targets"))))
stopifnot(identical(pitf, democracyData::pitf))
```

# Polyarchy

```{r polyarchy, results = 'asis'}
cite_dataset("polyarchy")

polyarchy_pmm <- extract_pmm_var(pmm_replication, pmm_polyarchy,
                                 include_in_output = include_in_output)

polyarchy <- redownload_polyarchy_original(verbose = verbose,
                                           include_in_output = include_in_output)

polyarchy_dimensions <- redownload_polyarchy_dimensions(verbose = verbose,
                                                        include_in_output = include_in_output)

stopifnot(identical(polyarchy_pmm, tar_read(polyarchy_pmm, store = here::here("_targets"))))
stopifnot(identical(polyarchy_pmm, democracyData::polyarchy_pmm))
stopifnot(identical(polyarchy, tar_read(polyarchy, store = here::here("_targets"))))
stopifnot(identical(polyarchy, democracyData::polyarchy))
stopifnot(identical(polyarchy_dimensions, tar_read(polyarchy_dimensions, store = here::here("_targets"))))
stopifnot(identical(polyarchy_dimensions, democracyData::polyarchy_dimensions))
```

# Political Regime Change (PRC)/Gasiorowski dataset

```{r prc_gasiorowski, results = 'asis'}
cite_dataset("prc_gasiorowski")

prc_gasiorowski <- prepare_prc(path = here::here("data-raw/Gasiorowski.csv"),
                               verbose = verbose,
                               include_in_output = include_in_output)

prc_pmm <- extract_pmm_var(pmm_replication, pmm_prc,
                           include_in_output = include_in_output)

stopifnot(identical(prc_gasiorowski, tar_read(prc_gasiorowski, store = here::here("_targets"))))
stopifnot(identical(prc_gasiorowski, democracyData::prc_gasiorowski))
stopifnot(identical(prc_pmm, tar_read(prc_pmm, store = here::here("_targets"))))
stopifnot(identical(prc_pmm, democracyData::prc_pmm))
```

# PIPE

```{r PIPE, results = 'asis'}
cite_dataset("PIPE")

# PIPE <- redownload_pipe(verbose = verbose,
#                         include_in_output = include_in_output)

```

# REIGN

```{r reign, results = 'asis'}
cite_dataset("reign")

REIGN <- redownload_reign(verbose = verbose,
                          include_in_output = include_in_output)

stopifnot(identical(REIGN, tar_read(REIGN, store = here::here("_targets"))))
stopifnot(identical(REIGN, democracyData::REIGN))
```

# SVMDI

```{r svmdi, results = 'asis'}
cite_dataset("svmdi")

svmdi <- redownload_svmdi(verbose = verbose,
                          include_in_output = include_in_output)

svmdi_2016 <- redownload_svmdi(release_year = 2016, verbose = verbose,
                               include_in_output = include_in_output)

stopifnot(identical(svmdi, tar_read(svmdi, store = here::here("_targets"))))
stopifnot(identical(svmdi, democracyData::svmdi))
stopifnot(identical(svmdi_2016, tar_read(svmdi_2016)))
stopifnot(identical(svmdi_2016, democracyData::svmdi_2016))
```
# Svolik

```{r svolik, results = 'asis'}
cite_dataset("svolik")

svolik_regime <- prepare_svolik_regime(path = here::here("data-raw/regime and no authority spells, country-year, 1946-2008.dta"),
                                       verbose = verbose,
                                       include_in_output = include_in_output)

stopifnot(identical(svolik_regime, tar_read(svolik_regime, store = here::here("_targets"))))
stopifnot(identical(svolik_regime, democracyData::svolik_regime))
```

# UDS

```{r uds, results = 'asis'}
cite_dataset("uds")

# The website is no longer online.

# uds_2014 <- redownload_uds(verbose = verbose,
#                            include_in_output = include_in_output)
# 
# uds_2011 <- redownload_uds(2011,
#                            verbose = verbose,
#                            include_in_output = include_in_output)
# 
# uds_2010 <- redownload_uds(2010,
#                            verbose = verbose,
#                            include_in_output = include_in_output)

# usethis::use_data(uds_2014, uds_2011, uds_2010, overwrite = TRUE)

cite_dataset("extended_uds")

set.seed(tar_meta() |> filter(name == "extended_uds") |> pull(seed))
extended_uds <- generate_extended_uds(verbose = verbose)

stopifnot(identical(extended_uds, tar_read(extended_uds, store = here::here("_targets"))))
stopifnot(identical(extended_uds, democracyData::extended_uds))
```

# Ulfelder

```{r ulfelder, results = 'asis'}
cite_dataset("ulfelder")

ulfelder <- redownload_ulfelder(verbose = verbose,
                                include_in_output = include_in_output)

ulfelder_extended <- redownload_ulfelder(verbose = verbose,
                                include_in_output = include_in_output,
                                extend = TRUE)

stopifnot(identical(ulfelder, tar_read(ulfelder, store = here::here("_targets"))))
stopifnot(identical(ulfelder, democracyData::ulfelder))
stopifnot(identical(ulfelder_extended, tar_read(ulfelder_extended, store = here::here("_targets"))))
stopifnot(identical(ulfelder_extended, democracyData::ulfelder_extended))
```

# UTIP

```{r utip, results = 'asis'}
cite_dataset("utip")

utip <- redownload_utip(verbose = verbose,
                        include_in_output = include_in_output)

stopifnot(identical(utip, tar_read(utip, store = here::here("_targets"))))
stopifnot(identical(utip, democracyData::utip))
```

# Vanhanen

```{r vanhanen, results = 'asis'}
cite_dataset("vanhanen")

vanhanen_pmm <- extract_pmm_var(pmm_replication, pmm_vanhanen,
                                verbose = verbose, include_in_output = include_in_output)

vanhanen <- prepare_vanhanen(path = here::here("data-raw/FSD1289/FSD1289/Study/data/daF1289e.csv"),
                             verbose = verbose, include_in_output = include_in_output)

stopifnot(identical(vanhanen_pmm, tar_read(vanhanen_pmm, store = here::here("_targets"))))
stopifnot(identical(vanhanen_pmm, democracyData::vanhanen_pmm))
stopifnot(identical(vanhanen, tar_read(vanhanen, store = here::here("_targets"))))
stopifnot(identical(vanhanen, democracyData::vanhanen))
```

# VDem

```{r, results='markup'}
cite_dataset("vdem")

vdem_simple <- prepare_vdem_simple(verbose = verbose, version = "13.0", include_in_output = include_in_output)

stopifnot(identical(vdem_simple, tar_read(vdem_simple, store = here::here("_targets"))))
stopifnot(identical(vdem_simple, democracyData::vdem_simple))
```


# Wahman Teorell and Hadenius

```{r wahman_teorell, results = 'asis'}
cite_dataset("wahman_teorell")

wahman_teorell_hadenius <- redownload_wahman_teorell_hadenius(verbose = verbose,
                                                              include_in_output = include_in_output)

stopifnot(identical(wahman_teorell_hadenius, tar_read(wahman_teorell_hadenius, store = here::here("_targets"))))
stopifnot(identical(wahman_teorell_hadenius, democracyData::wahman_teorell_hadenius))
```


