---
title: "Replicating and Extending the UD scores of Pemstein, Meserve, and Melton"
author: "Xavier Marquez"
date: "`r Sys.Date()`"
bibliography: bibfile.bib
biblio-style: american-political-science-review.csl
---

We can use this package to replicate and extend the [Unified Democracy Scores](http://www.unified-democracy-scores.org/) of @pmm2010uds2010 (which are no longer being updated or maintained), and in general to calculate latent variable indexes of democracy.[^workingPaper] This article is a modified version of the vignette for my package [QuickUDS](https://xmarquez.github.io/QuickUDS), which I am no longer actively maintaining; I am slowly migrating the functions in that package to this package to avoid having to update two different data sets of democracy measures.

[^workingPaper]: For more detail on the models used to generate these indexes, and their characteristics, see my working paper [@Marquez2016], available at [https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2753830](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2753830). 

You will need the package [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html) [@Chalmers2012], which can quickly compute full-information factor analyses. 

The basic procedure for replicating or extending the UD scores is very simple. 

1. Generate a dataset of democracy scores with a call to `generate_democracy_scores_dataset()`; 
2. Prepare the democracy data using the convenience function `prepare_democracy_data()`; 
3. Fit a simple [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html) model; 
4. Extract the calculated scores with a call to `democracy_scores()` or to `mirt::fscores()`. 

```{r, include=FALSE}
library(knitr)
opts_chunk$set(echo=TRUE)
```

## Preparing your democracy measures

The first step is to prepare the democracy measures for use with [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html). I focus first on replicating the 2011 release of the UDS, and then explain how to extend and augment these scores.

In order to replicate the original UD scores, we need to use PMM's replication dataset [@pmm2013uds2010]. This dataset is included this package: we just need to generate a data frame of democracy scores from all the datasets with names ending in `_pmm`. We can then use the function `prepare_democracy_data()` to put the data in the right format for use with [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html).

```{r}

library(mirt)
library(tidyverse)
library(democracyData)

identifiers <- c("extended_country_name", "GWn", "cown", "in_GW_system", "year")

democracy_data <- generate_democracy_scores_dataset(selection = "_pmm", include_extra_pmm = TRUE,
                                                    output_format = "wide") 
```

Before transformation by `prepare_democracy_data()`, the data looks like this:

```{r}
skimr::skim(democracy_data |> select(matches("pmm")))
```
After transformation, it looks like this:

```{r}
democracy_data_transformed <- prepare_democracy_data(democracy_data)

skimr::skim(democracy_data_transformed |> select(matches("pmm")))

```

The function `prepare_democracy_data()` gets rid of "empty rows" (country-years that have no measurements of democracy for the chosen indexes; such patterns will make [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html) fail) and transforms selected democracy indexes into ordinal variables suitable for use with [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html), mostly following the advice in Pemstein, Meserve, and Melton's original article (2010). 

In particular, `prepare_democracy_data()` will try to do the following on your dataset: 

* If a selected index contains the string `arat`, the function assumes the index is Arat's [@Arat1991] 0-109 democracy score, and cuts it into 7 intervals with the following cutoffs: 50, 60, 70, 80, 90, and 100. The resulting score is ordinal from 1 to 8 (following Pemstein, Meserve, and Melton's advice).

* If a selected index contains the string `bollen`, the function assumes the index is Bollen's [@Bollen2001] 0-100 democracy score, and cuts it into 10 intervals with the following cutoffs: 10,20,30,40,50,60,70,80, and 90. The resulting score is ordinal from 1 to 10 (following Pemstein, Meserve, and Melton's advice).

* If a selected index contains the string `wgi`, the function assumes the index is the World Governance Indicator's "Voice and Accountability" index [@wgi2017], and it will cut it into 20 categories. The resulting score is ordinal from 1 to 20.

* If a selected index contains the string `eiu`, the function assumes the index is the Economist Intelligence Unit's democracy index [@eiu2021], and it will cut it into 20 categories. The resulting score is ordinal from 1 to 20.

* If a selected index contains the string `hadenius_pmm`, the function assumes the index is Hadenius's 0-10 democracy score [@hadenius1992], and it will cut it into 8 intervals with the following cutoffs: 1, 2,3,4, 7, 8, and 9. The resulting score is ordinal from 1 to 8 (following Pemstein, Meserve, and Melton's advice). 

* If the selected index contains the string `munck`, the function assumes the index is Munck's 0-1 democracy score [@munck2009], and it will cut it into 4 intervals with the following cutoffs: 0.5,0.5,0.75, and 0.99. The resulting score is ordinal from 1 to 4 (following Pemstein, Meserve, and Melton's advice).

* If the selected index contains the string `peps`, the function assumes the index is one of the variants of the Participation-Enhanced Polity Score [@peps2006], and it will round its value (eliminating the decimal) and then transform it into an ordinal measure from 1 to 21.

* If the selected index contains the string `polity`, the function assumes this is the Polity IV or Polity 5 score [@polity2019; @polity2020], and it will thus set any values below -10 to NA and then transform the variable into an ordinal measure from 1 to 21.

* If the selected index contains the string `polyarchy_inclusion_dimension` or `polyarchy_contestation_dimension`, the function assumes this is one of the two dimensions of polyarchy estimated by @polyarchy_dimensions2008, and it will cut it into 20 categories. The resulting score is ordinal from 1 to 20.

* If the selected index contains the string `v2x`, the function assumes this is one of the v2x_ continuous indexes of democracy from the V-Dem dataset [@vdem11dataset], and it will cut it into 20 categories. The resulting score is ordinal from 1 to 20.

* If the selected index contains the string `csvdmi` or `svdmi_2016`, the function assumes this is one of the continuous indexes of democracy from the SVMDI dataset [@svmdi2016; @svmdi2018], and it will cut it into 20 categories. The resulting score is ordinal from 1 to 20.

* If the selected index contains the string `bti`, the function assumes this is the Bertelsman Transformation Index [@bti2020], and it will cut it into 20 categories. The resulting score is ordinal from 1 to 20.

* If the selected index contains the string `vanhanen_democratization` or `vanhanen_pmm`, the function assumes this is Vanhanen's index of democratization [@vanhanen2014], and it will cut it into 8 intervals with the following cutoffs: 5,10,15,20,25,30, and 35 (following Pemstein, Meserve, and Melton's advice). The resulting score is ordinal from 1 to 8.

`prepare_democracy_data()` will also work on column names that contain the following strings: 

* `anckar` [assumes it's the democracy indicator from @AnckarFredriksson2018]
* `anrr` [assumes it's the democracy indicator from @anrr2019]
* `blm` [assumes it's from @blm2005]
* `bmr` [assumes it's from @bmr2007]
* `doorenspleet` [assumes it's from @doorenspleet2000]
* `dsvmdi` [assumes it's the discrete machine-learning index @svmdi2018]
* `e_v2x` [assumes it's one of the "ordinal" indexes from the V-dem project, @vdem11dataset]
* `fh` or `freedomhouse` [assumes it's from @fh2020]
* `gwf` [assumes it's from @gwf2014 - the dichotomous democracy indicator only]
* `kailitz` [assumes it's from from @Kailitz2013 - democracy/electoral autocracy/non-democracy indicator only]
* `lied` or `lexical_index` [assumes it's from @LIED2015]
* `mainwaring` [assumes it's from @mainwaring2014]
* `magaloni` [assumes it's from @MagaloniChuMin2013]
* `pacl` or `cgv` [assumes it's from @pacl2010 or its later updates]
* `pitf` [assumes it's the measure of democracy used in @pitf2010; @pitf2015]
* `polyarchy` [assumes it's from @polyarchy1990]
* `prc` [assumes it's from @prc_gasiorowski1996 or its later update]
* `przeworski` [assumes it's the "regime" variable from @PIPE2013]
* `reign` [assumes it's the democracy/dictatorship indicator from @reign2016]
* `svolik` [assumes it's the democracy/dictatorship indicator from @svolik2012]
* `ulfelder` [assumes it's from @ulfelder2012]
* `utip` [assumes it's from @utip2008]
* `wahman_teorell_hadenius` or `wth` [assumes it's a democracy/non-democracy indicator from @wahman_teorell_hadenius2013]. 

In each of these cases the function `prepare_democracy_data()` transforms the values of the scores by running `as.numeric(unclass(factor(x)))`, which transforms each index into ordinal variables from 1 to the number of categories.

If you are using democracy indexes not included in the `democracy` dataset, or want to use your own custom measures of democracy, or transform them in a very particular way, you simply need to ensure that there are no "blank" country-years in your data (i.e., country-years without any democracy measurements; the package provides the convenience function `remove_empty_rows()` for this purpose) and that the indexes you are using are ordinal measures from 1 to N with every category present in the data. [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html) is pretty flexible and forgiving: it will accept ordinal variables in any range and will attempt to transform your indexes so that every category is within a distance of 1 of its neighboring categories. But it's useful to have a good sense of what the algorithm is doing to your data before you begin!

## Fitting a democracy model

After you've prepared the data, you can then fit a model as follows:

```{r}

replication_2011_model <- mirt(democracy_data_transformed |> select(matches("pmm")), model = 1, 
                               itemtype = "graded", SE = TRUE, 
                               verbose = FALSE)
                               
```

This just tells [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html) to fit a one-factor, full information graded response model like that in @pmm2010uds2010, and to calculate the standard errors for the coefficients. (See `?mirt` for details of the many options you can use to tweak your model, and see [my paper](https://papers.ssrn.com/sol3/papers.cfm?abstract_id=2753830) for a fuller description of why this model is useful here). 

Fitting this model is reasonably fast:

```{r}
replication_2011_model@time
```

We can easily check that this model converges and that it accounts for most of the variance in the democracy indexes:

```{r}
replication_2011_model
summary(replication_2011_model)
```

And we can then extract the latent democracy scores, either via `mirt::fscore()`, or via this package's convenient wrapper `democracy_scores` (which returns a tidy dataset with the latent scores and automatically calculates 95\% confidence intervals):

```{r}
replication_2011_scores <-  fscores(replication_2011_model, 
                                    full.scores = TRUE, 
                                    full.scores.SE = TRUE)
# Not a data frame, no country-years:
str(replication_2011_scores)

replication_2011_scores <- democracy_scores(model = replication_2011_model)

replication_2011_scores <- bind_cols(democracy_data, replication_2011_scores)

# A data frame with confidence intervals and country-years:

replication_2011_scores


```

We can check that these scores are, in fact, almost perfectly correlated with Pemstein, Meserve, and Melton's 2011 UDS release:

```{r}

uds <- generate_democracy_scores_dataset(selection = "uds", output_format = "wide")

replication_2011_scores <- replication_2011_scores |> 
  left_join(uds)

cor(replication_2011_scores |> select(matches("uds_2011"), z1), use = "pairwise")

```

(For more details on the relationship between the original UD scores and the replicated scores produced using this method, see my [working paper](http://ssrn.com/abstract=2753830)).

## Extending the model

Now suppose you want to create a new latent score derived but want to include other measures, or updated measures, or want to restrict your sources to dichotomous indicators of democracy or a particular set of measures that seem especially reliable. 

For example, suppose we want to use:

* The dichotomous indicator of democracy, adjusted for female suffrage, in version 3.0 of the Boix, Miller and Rosato dataset of political regimes [@bmr2007]
* The full extent of the Political Regime Change dataset [@prc_gasiorowski2002; @prc_gasiorowski1996], Vanhanen's index of democratization [@vanhanen2014], Bowman, Lehoucq, and Mahoney's data on Central America [@blm2005] and Mainwaring, Brinks and Perez-Linan's data on Latin America [@mainwaring2014], all of which go back to the beginning of the 20th century or before but are not used to their fullest extent in the official UD releases.
* One of the new V-Dem indexes of democracy, ordinal or continuous [@vdem11dataset]
* Renske Doorenspleet's dichotomous indicator of democracy including suffrage info [@doorenspleet2000]
* The World Governance Indicator's latest Voice and Accountability index 
* The most current release of Freedom House's data, to 2020, and the most current Polity data, to 2018
* The indicators of democracy in various autocratic regime datasets [@gwf2014; @svolik2012; @Kailitz2013; @wahman_teorell_hadenius2013]
* The 7-level Lexical Index of Democracy and Autocracy [@LIED2015]
* Jay Ulfelder's dichotomous indicator of democracy [@ulfelder2012]

This package makes the process extremely simple:

```{r}

all_dem <- generate_democracy_scores_dataset(output_format = "wide",
                                             verbose = FALSE)


other_dem <- all_dem |>
  select(any_of(identifiers), arat, blm, bmr_democracy_femalesuffrage,
         pmm_bollen, doorenspleet, wgi_democracy, fh_total_reversed, 
         gwf_democracy_extended_strict, pmm_hadenius, kailitz_tri, svolik_democracy, 
         lexical_index, ulfelder_democracy_extended, prc, mainwaring, 
         vanhanen_democratization, v2x_polyarchy)

other_dem <- prepare_democracy_data(other_dem)

extended_model <- mirt(other_dem |> select(-any_of(identifiers)), 
                       model = 1, itemtype = "graded", SE = TRUE, verbose = FALSE)

summary(extended_model)

extended_scores <- democracy_scores(model = extended_model)

extended_scores <- bind_cols(other_dem |> select(any_of(identifiers)),
                             extended_scores)

extended_scores <- extended_scores |>
  left_join(uds |> select(any_of(identifiers), matches("_mean")))
```

[`mirt`](https://cran.r-project.org/web/packages/mirt/index.html) will stop by default after 500 EM cycles, but some models will take longer to converge. If your model has not converged after 500 iterations of the algorithm, you can try increasing the number of cycles with the `technical` option. Use `?mirt` for more details.

One important point to note about latent variable democracy scores is that they are normalized with mean zero and standard deviation one, so a score of 1 just means that the country-year is 1 standard deviation more democratic than the average country-year in the sample. But this means that adding extra country-years to our model will typically result in scores that have a higher mean (though usually smaller standard errors) than the original UD model, given that the world has become substantially more democratic over the last two centuries:

```{r, fig.height=8, fig.width=7}
countries <- c("United States of America", 
               "United Kingdom","Argentina",
               "Chile","Venezuela","Spain")

data <- extended_scores |> 
  filter(extended_country_name %in% countries) |>
  tidyr::gather(measure, zscore, uds_2014_mean, z1) |>
  filter(!is.na(zscore), year >=1946, year < 2008) |>
  mutate(measure = case_when(
    measure == "uds_2010_mean" ~ "2010 release of UDS",
    measure == "uds_2011_mean" ~ "2011 release of UDS",
    measure == "uds_2014_mean" ~ "2014 release of UDS",
    measure == "z1_matched" ~ "Extended replication score",
    TRUE ~ measure
    )
)

ggplot(data = data, 
       aes(x = year, y = zscore, color = measure)) + 
  geom_path() + 
  theme_bw() + 
  labs(x = "Year", y = "Latent unified democracy scores,\nper year")  + 
  theme(legend.position="bottom") + 
  guides(color = guide_legend(ncol = 1),fill = guide_legend(nrow = 1)) + 
  facet_wrap(~extended_country_name, ncol = 2)  

```

If necessary, one can therefore "match" the extended scores to the official UD release by substracting the mean of the extended scores for the period of the UD release one wants to match from the extended scores (that is, making the mean of the extended scores equal to zero for the period of adjustment):

```{r, fig.height=8, fig.width=7}

matched_data <- extended_scores |> 
  filter(!is.na(uds_2014_mean)) |>
  mutate(z1_matched = z1 - mean(z1, na.rm = TRUE), 
         z1_pct975_matched = z1_pct975 - mean(z1, na.rm = TRUE), 
         z1_pct025_matched = z1_pct025 - mean(z1, na.rm = TRUE))

matched_data <- matched_data |> 
  filter(extended_country_name %in% countries) |>
  tidyr::gather(measure, zscore, uds_2014_mean, z1_matched) |>
  filter(!is.na(zscore), year >=1946, year < 2008) |>
  mutate(measure = case_when(
    measure == "uds_2010_mean" ~ "2010 release of UDS",
    measure == "uds_2011_mean" ~ "2011 release of UDS",
    measure == "uds_2014_mean" ~ "2014 release of UDS",
    measure == "z1_matched" ~ "Matched extended replication score",
    TRUE ~ measure
    ))

ggplot(data = matched_data, 
       aes(x = year, y = zscore, color = measure)) + 
  geom_path() + 
  theme_bw() + 
  labs(x = "Year", y = "Latent unified democracy scores,\nper year")  + 
  theme(legend.position="bottom") + 
  guides(color = guide_legend(ncol=1),fill = guide_legend(nrow=1)) + 
  facet_wrap(~extended_country_name,ncol=2)  

```

In the graph above, we can see that the 2014 release of the UDS seems to overestimate the degree of democracy in the USA in the early decades of 1950 relative to the "extended" scores. 

These scores have a more natural interpretation when transformed to a 0-1 index using the cumulative distribution function as the "probability that a country-year is democratic" (so the 0 is now a natural minimum, and 1 a natural maximum). These indexes are automatically produced by the function `democracy_scores`; they are in the column `z1_as_prob` of the output, and are produced by applying the `pnorm` function to `z1`, as follows:

```{r}
extended_scores <- extended_scores |> 
  mutate(index = pnorm(z1), 
         index_pct025 = pnorm(z1_pct025), 
         index_pct975 = pnorm(z1_pct975))

# These are equal to z1_as_prob, which is automatically calculated:

extended_scores |> filter(index != z1_as_prob)
```

It is also possible to to set the cutpoint for this index at, for example, the average cutpoint in the latent variable of the dichotomous indexes of democracy (so that 0.5 correponds more naturally to the point at which a regime could be either democratic or non-democratic according to the dichotomous measures of democracy included in your model). These scores are also automatically calculated (they are in the column `z1_adj`) but they can also be manually added as follows: 

```{r, fig.height=8, fig.width=7}
cutpoints_extended <- cutpoints(extended_model)

cutpoints_extended

dichotomous_cutpoints <- cutpoints_extended |> 
  group_by(variable) |>
  filter(n() == 1) 

dichotomous_cutpoints

avg_dichotomous <- mean(dichotomous_cutpoints$estimate)

avg_dichotomous

extended_scores <- extended_scores |> mutate(adj_z1 = z1 - avg_dichotomous, 
                                        adj_pct025 = z1_pct025 - avg_dichotomous, 
                                        adj_pct975 =z1_pct975 - avg_dichotomous,
                                        index = pnorm(adj_z1),
                                        index_pct025 = pnorm(adj_pct025),
                                        index_pct975 = pnorm(adj_pct975))

ggplot(data = extended_scores |> filter(extended_country_name %in% countries), 
       aes(x= year, y = index, 
           ymin = index_pct025, ymax = index_pct975)) + 
  geom_line() + 
  geom_ribbon(alpha=0.2) + 
  theme_bw() + 
  labs(x = "Year", y = "Latent unified democracy scores,\nper year\nconverted to 0-1 probability scale")  + 
  theme(legend.position="bottom") + 
  guides(color = guide_legend(ncol=1),fill = guide_legend(nrow=1)) + 
  geom_hline(yintercept=0.5,color="red") +
  facet_wrap(~extended_country_name,ncol=2)  
```

A pre-computed and documented version of the extended UDS scores, with data from all the indexes mentioned above, plus the participation-enhanced Polity Scores of @peps2006, a trichotomous democracy indicator calculated from Magaloni, Min, and Chu's "Autocracies of the World" datset [@MagaloniChuMin2013], a dichotomous democracy indicator calculated from @utip2008, the REIGN dataset of @reign2016, which extends @gwf2014, a dichotomous democracy indicator from @anrr2019, the Bertelsmann Transformation index [@bti2020], and an indicator of democracy used by the Political Instability Task Force [@pitf2010; @pitf2015], is included with the package; it can be loaded by simply typing `extended_uds`. Use `?extended_uds` to examine the documentation for all its variables, and see my working paper ([Marquez 2016](http://ssrn.com/abstract=2753830)) for more detail on the data and its uses.

The function `generate_extended_uds()` recreates these scores in one line of code, at the cost of some flexibility. It requires that the vdem package be installed; you can install it by using `remotes::install_github("xmarquez/vdem)`.

## Other Extensions

We can also use this method to create indexes from specific types of scores, such as dichotomous measures of democracy. Here we compute a 2-parameter logistic model from all dichotomous indexes of democracy (excluding near-duplicates):

```{r}
dichotomous_dem <- all_dem |>
  select(any_of(identifiers), where(~n_distinct(.) <= 3))  |>
  select(-pacl, 
         -bmr_democracy_omitteddata, -bmr_democracy, 
         -wth_democ1, 
         -gwf_democracy_extended, -utip_dichotomous)

dichotomous_dem <- prepare_democracy_data(dichotomous_dem)

dichotomous_model <- mirt(dichotomous_dem |> select(-any_of(identifiers)),
                          model = 1, itemtype = "2PL", SE = TRUE, verbose = FALSE)

summary(dichotomous_model)

dichotomous_scores <- democracy_scores(dichotomous_model)

dichotomous_scores <- bind_cols(dichotomous_dem |> select(any_of(identifiers)),
                                dichotomous_scores)


ggplot(data = dichotomous_scores |> filter(extended_country_name %in% countries), 
       aes(x= year, y = z1_as_prob, 
           ymin = z1_pct025_as_prob, ymax = z1_pct975_as_prob)) + 
  geom_line() + 
  geom_ribbon(alpha=0.2) + 
  theme_bw() + 
  labs(x = "Year", y = "Latent unified democracy scores,\nper year\nconverted to 0-1 probability scale")  + 
  theme(legend.position="bottom") + 
  guides(color = guide_legend(ncol=1),fill = guide_legend(nrow=1)) + 
  geom_hline(yintercept=0.5,color="red") +
  facet_wrap(~extended_country_name,ncol=2)
```

As @svmdi2021 note, latent variable indexes suffer from arbitrary changes in level related to variables entering into or out of the source data. One way to get around this is to use a panel, with every measure present for every country-year in the panel. For example, suppose we're interested only in measures with long coverage. Here we select a set of indexes with coverage down to the 19th century and then select the set of rows for which all measures exist, producing a panel with 159 countries and scores from 1919 to 2003.

```{r}
full_panel <- all_dem |>
  select(any_of(identifiers), reign_democracy, polity2, 
         bmr_democracy_femalesuffrage, v2x_polyarchy,
         ulfelder_democracy_extended, bnr_extended, 
         magaloni_democracy_extended, csvmdi, pitf,
         anckar_democracy, PEPS1v, vanhanen_democratization) |>
  rowwise() |>
  mutate(num_nas = sum(is.na(c_across(-any_of(identifiers))))) |>
  filter(num_nas == 0) |>
  ungroup() |>
  select(-num_nas)

full_panel <- prepare_democracy_data(full_panel)

panel_model <- mirt(full_panel |> select(-any_of(identifiers)),
                    model = 1, itemtype = "graded", SE = TRUE, 
                    verbose = FALSE, technical = list(NCYCLES = 1000))

panel_model@time

summary(panel_model)

panel_scores <- democracy_scores(panel_model)

panel_scores <- bind_cols(full_panel |> select(any_of(identifiers)),
                                panel_scores)

skimr::skim(panel_scores)

ggplot(data = panel_scores |> filter(extended_country_name %in% countries), 
       aes(x= year, y = z1_as_prob, 
           ymin = z1_pct025_as_prob, ymax = z1_pct975_as_prob)) + 
  geom_line() + 
  geom_ribbon(alpha=0.2) + 
  theme_bw() + 
  labs(x = "Year", y = "Latent unified democracy scores,\nper year\nconverted to 0-1 probability scale")  + 
  theme(legend.position="bottom") + 
  guides(color = guide_legend(ncol=1),fill = guide_legend(nrow=1)) + 
  geom_hline(yintercept=0.5,color="red") +
  facet_wrap(~extended_country_name,ncol=2)

```

Or suppose we're interested in a particular coverage period, including only measures that have data to 2018:

```{r}
full_panel <- all_dem |>
  pivot_longer(-any_of(identifiers), values_drop_na = TRUE) |>
  filter(name %in% name[year == 2018]) |>
  filter(year <= 2018) |>
  pivot_wider(id_cols = any_of(identifiers), names_from = "name", values_from = "value") |>
  unnest(fh_total_reversed:eiu) |>
  select(-pitf_binary, -dsvmdi, -polityIV, -polity2IV, 
         -polity, -bti_democracy, -vanhanen_competition, 
         -vanhanen_participation) |>
  rowwise() |>
  mutate(num_nas = sum(is.na(c_across(-any_of(identifiers))))) |>
  filter(num_nas == 0) |>
  ungroup() |>
  select(-num_nas)

full_panel <- prepare_democracy_data(full_panel)

panel_model <- mirt(full_panel |> select(-any_of(identifiers)),
                    model = 1, itemtype = "graded", SE = TRUE, 
                    verbose = FALSE, technical = list(NCYCLES = 1000))

panel_model@time

summary(panel_model)

panel_scores <- democracy_scores(panel_model)

panel_scores <- bind_cols(full_panel |> select(any_of(identifiers)),
                                panel_scores)

skimr::skim(panel_scores)

ggplot(data = panel_scores |> filter(extended_country_name %in% countries), 
       aes(x= year, y = z1_as_prob, 
           ymin = z1_pct025_as_prob, ymax = z1_pct975_as_prob)) + 
  geom_line() + 
  geom_ribbon(alpha=0.2) + 
  theme_bw() + 
  labs(x = "Year", y = "Latent unified democracy scores,\nper year\nconverted to 0-1 probability scale")  + 
  theme(legend.position="bottom") + 
  guides(color = guide_legend(ncol=1),fill = guide_legend(nrow=1)) + 
  geom_hline(yintercept=0.5,color="red") +
  facet_wrap(~extended_country_name,ncol=2)

```

## Extracting useful information from a model

The [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html) package offers a great number of powerful tools to examine and diagnose the fitted model, including functions to extract model cutpoints and item information curves. But this package also contains two convenience functions that wrap [`mirt`](https://cran.r-project.org/web/packages/mirt/index.html) tools to quickly extract democracy rater discrimination parameters, rater cutoffs, and rater information curves from a model produced by this procedure in a tidy data frame format suitable for graphing. Here, for example, we can replicate the figures in PMM's original paper:

```{r, fig.height=8, fig.width=7, warning= FALSE}
replication_2011_cutpoints <- cutpoints(replication_2011_model, type ="score")
replication_2011_cutpoints

# We plot the "normalized" cutpoints ("estimate," in the same scale as the latent scores), 
# not the untransformed ones ("par")

ggplot(data = replication_2011_cutpoints, 
       aes(x = variable, y = estimate, 
           ymin = pct025, ymax = pct975))  + 
  theme_bw() + 
  labs(x="",y="Unified democracy level rater cutoffs") + 
  geom_point() + 
  geom_errorbar() + 
  geom_hline(yintercept =0, color="red") + 
  coord_flip()

# We can also plot discrimination parameters, which are in a different scale:
replication_2011_discrimination <- cutpoints(replication_2011_model, 
                                             type ="discrimination")

replication_2011_discrimination

ggplot(data = replication_2011_discrimination, 
       aes(x=reorder(variable,estimate),
           y = estimate, ymin = pct025, 
           ymax = pct975))  + 
  theme_bw() + 
  labs(x="",y="Discrimination parameter for each rater
       \n(higher value means fewer idiosyncratic\nerrors relative to latent score)") + 
  geom_point() + 
  geom_errorbar() + 
  coord_flip()

# And we can plot item information curves for each rater:
replication_2011_info <- raterinfo(replication_2011_model)
replication_2011_info

ggplot(data = replication_2011_info, aes(x=theta,y=info)) + 
  geom_path() + 
  facet_wrap(~rater) + 
  theme_bw() + 
  labs(x="Latent democracy score",y = "Information") + 
  theme(legend.position="bottom")

```

Finally, the package offers a simple function to estimate the probability that a given country is more democratic than another in a given year, accounting for the uncertainty in the UD-style measures. For example, suppose we want to know the probability that the USA was more democratic than France in the year 2000 for both the replicated 2011 scores and our extended model:

```{r}
prob_more(replication_2011_scores, "United States of America","France", 2000)
prob_more(extended_scores, "United States of America","France", 2000)

```

Or perhaps we wish to know the probability that the United States was more democratic in the year 2000 than in the year 1953:

```{r}
prob_more(replication_2011_scores, 
          "United States of America",
          "United States of America", 
          c(2000,1953))
prob_more(extended_scores, 
          "United States of America",
          "United States of America", 
          c(2000,1953))

```


# References

