---
title: "Understanding the geographic and temporal coverage of existing democracy indexes"
author: "Xavier Marquez"
date: "`r format(Sys.Date())`"
bibliography: bibfile.bib
csl: american-political-science-review.csl
format:
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    code-fold: true
    code-tools: true
    fig-cap-location: top
    tbl-cap-location: top
    smooth-scroll: true
    link-external-icon: true
execute:
  echo: false
  warning: false
  message: false
  cache: true
knitr:
  opts_chunk:
    collapse: true
    comment: "#>"
---

```{r}
#| label: setup
#| cache: false
library(democracyData)
library(tidyverse)

```

The geographical and temporal coverage of different democracy indexes varies greatly.

```{r}

theme_set(theme_bw())

long_data <- generate_democracy_scores_dataset(verbose = FALSE)

long_data <- long_data |>
  group_by(measure) |>
  mutate(value_rescaled = scales::rescale(value, to = c(0,1))) |>
  group_by(measure, extended_country_name) |>
  mutate(coverage = n()) |>
  ungroup() 


wars <- tibble(xmin = c(1775, 1789, 1914, 1939, 1989), 
                      xmax = c(1787, 1810, 1918, 1945, 1991),
                      ymin = -Inf, ymax = Inf)

state_system <- democracyData:::data |> 
  distinct(extended_country_name, GWn, extended_region, extended_continent) 

long_data <- left_join(long_data, state_system, relationship = "many-to-many")

long_data <- distinct(long_data)

duplicated <- long_data |> 
  group_by(extended_country_name, GWn, year, measure) |> 
  filter(n() > 1)

duplicated <- duplicated |>
  slice_tail(n = 1)

long_data <- anti_join(long_data, duplicated)
```

[VDem](../../reference/vdem_simple.html) and [LIED](../../reference/LIED.html) code the largest total number of country-years, while specialist indexes like [BLM](../../reference/blm.html) (which covers only 5 countries in Latin America) code the smallest number. Given the increase in the number of states in the state system, it's no surprise that coverage increases greatly after WWII, with many datasets covering a large number of countries. 

```{r}
#| label: fig-temporal-coverage-datasets
#| fig-cap: "Temporal coverage of each dataset, ordered by total number of country-years coded by all measures in the dataset (some datasets offer more than one measure of democracy). Shaded red bands mark the American and French revolutions, the First and Second World Wars, and the end of the Cold War."
#| fig-height: 12
#| warning: false
#| message: false

ggplot(long_data |>
         group_by(dataset, year) |>
         summarise(num_countries = n_distinct(extended_country_name)) |>
         filter(!is.na(num_countries))) +
  geom_tile(aes(x = year, y = fct_reorder(dataset, num_countries, sum), alpha = num_countries)) +
  labs(y = "", fill = "Number of countries", alpha = "Number of countries",
       title = NULL) +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  theme_bw() +
  theme(legend.position = "bottom")

```

```{r}
#| label: fig-temporal-coverage-measures
#| fig-cap: "Temporal coverage of every individual democracy measure, ordered by total number of country-years. Shaded red bands mark the American and French revolutions, the First and Second World Wars, and the end of the Cold War."
#| fig-height: 20
#| warning: false
#| message: false
ggplot() +
  geom_tile(data = long_data |>
  group_by(measure, year) |>
  summarise(num_countries = n_distinct(extended_country_name)) |>
  filter(!is.na(num_countries)),
  aes(x = year, y = (fct_reorder(measure, num_countries, sum)), alpha = num_countries)) +
  labs(y = "", alpha = "Number of countries",
       title = NULL) +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  theme(legend.position = "bottom")

```

As we can see, the majority of datasets measure democracy between the end of the Second World War and the first two decades after the end of the Cold War. The most heavily measured country is the United States; the earliest measurement is for Iran (1502, in the Ulfelder extended dataset; this is the regime start date). The vast majority of measurements are after 1945, and roughly a dozen datasets are currently maintained through 2024-2025, including LIED, V-Dem, VaPoReg, BMR, Anckar, Polity5, PACL_update, SVMDI, FH (frozen at the 2025 release), Svolik, EIU, BTI, and PITF.

```{r}
#| label: fig-coverage-by-country
#| fig-cap: "Coverage of democracy measures per country or territory, ordered by total number of democracy measurements. Alpha shading marks whether the unit is in the Gleditsch-Ward state system in that year."
#| fig-height: 30
#| warning: false
#| message: false

ggplot(data = long_data |>
  group_by(extended_country_name, year) |>
  summarise(
    num_measures = n(),
    in_GW_system = any(dplyr::coalesce(in_GW_system, FALSE)),
    .groups = "drop"
  ) |>
  filter(!is.na(num_measures)) |>
  group_by(extended_country_name) |>
  mutate(total_measures = sum(num_measures)) |>
  ungroup()) +
  geom_tile(aes(x = year, y = (fct_reorder(extended_country_name, total_measures, mean)),
             alpha = in_GW_system, fill = num_measures)) +
  labs(y = "", fill = "Number of measures", alpha = "In state system",
       title = NULL)  +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  scale_fill_gradient2(midpoint = 20, mid = "grey") +
  scale_alpha_discrete(range = c(0.3,1)) +
  theme_bw() +
  theme(legend.position = "bottom")

```

There can be a fair amount of disagreement in these measures, especially at regime transition points (see article on correlations between measures for more discussion). Consider the USA, Venezuela, and Russia (darker colors = more democratic):

```{r}
#| label: fig-coverage-usa
#| fig-cap: "Measurements of democracy for the United States of America across all indexes, rescaled to 0-1. Shaded red bands mark the American and French revolutions, the First and Second World Wars, and the end of the Cold War."
#| fig-height: 20
long_data |>
  filter(extended_country_name == "United States of America") |>
  ggplot() +
  geom_tile(aes(x = year, fill = value_rescaled, y = fct_reorder(measure, coverage))) +
  labs(fill = "Measure value\n(Rescaled to 0-1)",
       y = "",
       title = NULL)  +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  scale_fill_viridis_c(direction = -1)  +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-coverage-venezuela
#| fig-cap: "Measurements of democracy for Venezuela across all indexes, rescaled to 0-1. Shaded red bands mark the American and French revolutions, the First and Second World Wars, and the end of the Cold War."
#| fig-height: 20
long_data |>
  filter(extended_country_name == "Venezuela") |>
  ggplot() +
  geom_tile(aes(x = year, fill = value_rescaled, y = fct_reorder(measure, coverage))) +
  labs(fill = "Measure value\n(Rescaled to 0-1)",
       y = "",
       title = NULL)  +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  scale_fill_viridis_c(direction = -1) +
  theme(legend.position = "bottom")
```

```{r}
#| label: fig-coverage-russia
#| fig-cap: "Measurements of democracy for Russia / the Soviet Union across all indexes, rescaled to 0-1. Shaded red bands mark the American and French revolutions, the First and Second World Wars, and the end of the Cold War."
#| fig-height: 20
long_data |>
  filter(extended_country_name == "Russia (Soviet Union)") |>
  ggplot() +
  geom_tile(aes(x = year, fill = value_rescaled, y = fct_reorder(measure, coverage))) +
  labs(fill = "Measure value\n(Rescaled to 0-1)",
       y = "",
       title = NULL)  +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  scale_fill_viridis_c(direction = -1) +
  theme(legend.position = "bottom")
```

We can also take a look at the average level of democracy for each measure, across the world. As we can see, the world has become more democratic (darker colors = more democratic), though the particular moment at which this happened varies from measure to measure.

```{r}
#| label: fig-world-averages-by-dataset
#| fig-cap: "Average level of democracy for the world by year, by dataset (rescaled to 0-1, darker = more democratic). Alpha shading reflects the number of countries measured in each year."
#| fig-height: 12
long_data |>
  group_by(dataset, year) |>
  summarise(value_rescaled = mean(value_rescaled),
            coverage = sum(coverage),
            num_countries = n_distinct(extended_country_name)) |>
  ggplot() +
  geom_tile(aes(x = year, fill = value_rescaled, y = fct_reorder(dataset, coverage),
                alpha = num_countries)) +
  labs(fill = "Avg. score (0-1, 1 = democracy)",
       y = "",
       title = NULL,
       alpha = "Countries measured")  +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  scale_fill_viridis_c(direction = -1) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "vertical")

```

```{r}
#| label: fig-world-averages-by-measure
#| fig-cap: "Average level of democracy for the world by year, by individual measure (rescaled to 0-1, darker = more democratic). Alpha shading reflects the number of countries measured in each year."
#| message: false
#| warning: false
#| fig-height: 15
long_data |>
  group_by(measure, year) |>
  summarise(value_rescaled = mean(value_rescaled),
            coverage = sum(coverage),
            num_countries = n_distinct(extended_country_name)) |>
  ggplot() +
  geom_tile(aes(x = year, fill = value_rescaled, y = fct_reorder(measure, coverage),
                alpha = num_countries)) +
  labs(fill = "Avg. score (0-1, 1 = democracy)",
       y = "",
       title = NULL,
       alpha = "Countries measured")  +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  scale_fill_viridis_c(direction = -1) +
  theme_bw() +
  theme(legend.position = "bottom",
        legend.box = "vertical")

```

Finally, we visualize the average level of democracy per country (averaging across all measures that code a country-year; darker colors = more democratic).

```{r}
#| label: fig-average-level-by-country
#| fig-cap: "Average level of democracy per country, averaged across all measures that code a given country-year (rescaled to 0-1, darker = more democratic). Alpha shading reflects the number of measures coding each country-year."
#| message: false
#| warning: false
#| fig-height: 30
theme_set(theme_bw())
long_data |>
  group_by(extended_country_name, year) |>
  summarise(value_rescaled = mean(value_rescaled),
            coverage = sum(coverage),
            num_measures = n_distinct(measure)) |>
  ggplot() +
  geom_tile(aes(x = year, fill = value_rescaled, y = fct_reorder(extended_country_name, coverage),
                alpha = num_measures)) +
  labs(fill = "Avg. score (0-1, 1 = democracy)",
       y = "",
       title = NULL,
       alpha = "Number of measures")  +
  geom_rect(data = wars, aes(xmin = xmin,
                      xmax = xmax,
                      ymin = -Inf, ymax = Inf), fill = "red", alpha = 0.3) +
  scale_fill_viridis_c(direction = -1) +
  theme(legend.position = "bottom",
        legend.box = "vertical")

```


